---
editor_options:
  chunk_output_type: console
---

NOTES:

-   This file is *not* included in the [source package](https://r-pkgs.org/package-structure-state.html#source-package) because of the [`.nopurl` suffix in the
    filename](https://rpkg.dev/pkgpurl/reference/purl_rmd.html#-rmd-files-excluded-from-purling).

-   The chunks below have to be manually executed in order to regenerate the package data.

# Exported data

NOTES:

-   When used internally, other than internal data, exported data must always be referred to by its explicit namespace.

## `regex_text_normalization`

```{r}
regex_text_normalization <-
  tibble::tibble(id = character(),
                 category = character(),
                 purpose = character(),
                 pattern = list(),
                 replacement = character()) %>%
  # harmonize punctuation
  tibble::add_row(id = "uniform_quotation_marks",
                  category = "harmonize_punctuation",
                  purpose = "use typewriter double quotes (`\"`) as quotation marks",
                  pattern = list("[“”„‟⹂]"),
                  replacement = "\"") %>%
  tibble::add_row(id = "uniform_apostrophes",
                  category = "harmonize_punctuation",
                  purpose = "use typewriter single quotes (`'`) as apostrophes",
                  pattern = list("[’‘‚‛]"),
                  replacement = "'") %>%
  # prettify punctuation
  tibble::add_row(id = "no_break_percentages",
                  category = "prettify_punctuation",
                  purpose = "use narrow non-breaking space between numbers and percentage signs",
                  pattern = list("\\b(\\d+) ?(\\\\?%(\\W|$)|Prozent|percent|per\\scent)(?![\"'])"),
                  replacement = "\\1\u202f\\2") %>%
  tibble::add_row(id = "no_break_abbreviations_german",
                  category = "prettify_punctuation",
                  purpose = "use narrow non-breaking space between characters of common German abbreviations",
                  pattern = list("(?i)\\b(z\\.) ?(B\\.)",
                                 "(?i)\\b(z\\.) ?(T\\.)",
                                 "(?i)\\b(u\\.) ?(a\\.)",
                                 "(?i)\\b(d\\.) ?(h\\.)"),
                  replacement = "\\1\u202f\\2") %>%
  tibble::add_row(id = "no_break_equals_sign",
                  category = "prettify_punctuation",
                  purpose = "use narrow non-breaking space before and after certain assignments and equality comparisons (`=` and `==`)",
                  pattern = list("(?<= \\(\\w{1,32})(?: ?)(==?)(?: ?)(?=\\d+)"),
                  replacement = "\u202f\\1\u202f") %>%
  tibble::add_row(id = "en_dash_value_ranges",
                  category = "prettify_punctuation",
                  purpose = "use [en dash](https://www.thepunctuationguide.com/en-dash.html) instead of hyphen in value ranges",
                  pattern = list("(?<!-)(\\b\\d+)-(\\d+\\b)(?!-)"),
                  replacement = "\\1\u2013\\2")
```

## `regex_file_normalization`

Note that the ICU regex engine used by yay/stringr/stringi [doesn't yet support](https://unicode-org.atlassian.net/browse/ICU-11291) case-conversion replacement
syntax (`\U`, `\L` and friends).

```{r}
regex_file_normalization <-
  tibble::tibble(id = character(),
                 category = character(),
                 purpose = character(),
                 pattern = list(),
                 replacement = character()) %>%
  # ensure terminating newline
  tibble::add_row(id = "terminating_newline",
                  category = "posix",
                  purpose = "ensure file ends with the [POSIX-standard newline control character `LF`](https://en.wikipedia.org/wiki/Newline) (aka `\n`) ",
                  pattern = list("(.)(?!\\n)\\Z"),
                  replacement = "\\1\n") %>%
  # lint [Go template](https://golang.org/pkg/text/template/) syntax
  tibble::add_row(id = "go_template_opening_double_brace_space",
                  category = "go_template",
                  purpose = "ensure there's a space after opening double braces in Go templates",
                  pattern = list("\\{\\{(?!( |- |%|<|/\\*))"),
                  replacement = "{{ ") %>%
  tibble::add_row(id = "go_template_closing_double_brace_space",
                  category = "go_template",
                  purpose = "ensure there's a space before closing double braces in Go templates",
                  pattern = list("([^-%>/ ]|[^ ]-)\\}\\}"),
                  replacement = "\\1 }}")
```

## Write data

Save all of the above data objects to separate files under `data/*.rda`. Note that when documenting them, they mustn't be explicitly `@export`ed since they're
already automatically exported and thus available to package users.

```{r}
usethis::use_data(regex_text_normalization,
                  regex_file_normalization,
                  internal = FALSE,
                  overwrite = TRUE,
                  compress = "xz",
                  version = 3L)
```
