---
editor_options:
  chunk_output_type: console
---

# NOTES

-   This file is *not* included in the [source package](https://r-pkgs.org/structure.html#sec-source-package) because of the [`.nopurl` suffix in its
    filename](https://pkgpurl.rpkg.dev/reference/purl_rmd.html#-rmd-files-excluded-from-purling).

-   The chunks below have to be manually executed in order to regenerate the package data.

-   Although the datasets below are saved as ["internal data" in `R/sysdata.rda`](https://r-pkgs.org/data.html#sec-data-sysdata), they can still be exported and
    documented (by documenting the dataset's quoted name in the main `.Rmd` source file – which only works when the dataset is also `@export`ed), something [not
    explicitly mentioned](https://coolbutuseless.github.io/2018/12/10/r-packages-internal-and-external-data/) in the book [R
    Packages](https://r-pkgs.org/data.html#sec-data-data). To do so, you first need to manually add the `export()` directive in the NAMESPACE file since
    roxygen2 won't add it automatically.

# Define data

## `pkg_config`

```{r}
pkg_config <-
  tibble::tibble(key = character(),
                 default_value = list(),
                 default_value_dynamic = character(),
                 require = logical(),
                 description = character()) |>
  tibble::add_row(key = "netlify_token",
                  require = TRUE,
                  description = paste0("[Netlify personal access token](https://docs.netlify.com/api/get-started/#authentication) used for authentication ",
                                       "in the [`netlify_*()` functions](https://yay.rpkg.dev/reference/index.html#netlify).")) |>
  tibble::add_row(key = "porkbun_api_key",
                  require = TRUE,
                  description = paste0("[Porkbun API key](https://porkbun.com/api/json/v3/documentation#Authentication) used for authentication in the ",
                                       "[`porkbun_*()` functions](https://yay.rpkg.dev/reference/index.html#porkbun).")) |>
  tibble::add_row(key = "porkbun_secret_api_key",
                  require = TRUE,
                  description = paste0("[Porkbun secret API key](https://porkbun.com/api/json/v3/documentation#Authentication) used for authentication in ",
                                       "the [`porkbun_*()` functions](https://yay.rpkg.dev/reference/index.html#porkbun)."))
```

# Write data

Save all the small data objects as a single internal file `R/sysdata.rda`. Note that when documenting them, they must be explicitly `@export`ed to be available
to package users.

```{r}
usethis::use_data(pkg_config,
                  internal = TRUE,
                  overwrite = TRUE,
                  compress = "xz",
                  version = 3L)
```
